---
title: "IF H1 pt. 2"
author: "Rodrigo, Valerie, Marcela"
date: "26/8/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("gmodels")
library("ggplot2")
library("rpart.plot")
library("tinytex")
library("kableExtra")
library("tidyquant")
library("knitr")
library("plotly")
library("roptions")
library("fpp")
library("fpp2")
library("dplyr")
library("latexpdf")
```

# Abstract [TBU]

# Introduction [TBU]

# Part 1 [TBU]

# Creating a bull spread using call and put options for Tesla stock
As John C. Hall states on his well-known book, *Options, futures and other derivatives* (2015), a spread is a trading strategy involving taking positions in 2 or more options of the same type (i.e. calls / puts). Furthermore, we also learn that spreads can be either bullish or bearish, depending on whether the profit is achieved on a bullish or bearish market. In this section of the paper, **we focus on creating two bull spreads by taking opposing positions on call and put options for Tesla (TSLA) stock**. Additionally, both spreads we create are bullish, that is, we are assuming an increase in Tesla's stock price in order to make a profit.

The goal of this exercise is to demonstrate that it is possible to create a similar profit curve by building a spread using call and put options. In other words, we aim to show that both types of options can be used in order to create a profitable spread-trading strategy.

It is worth noting that, just as in section 1, we also leverage Yahoo Finance's database to retrieve the relevant variables for this exercise, such as the current stock (defined as *S0*) and options prices. Please note we take the *ask price* for long positions and *bid price* for short positions, as we are doing this exercise from a retail trader's perspective (and retail traders pay the dealer's ask price and get paid their bid price, since the retail trader does not have the ability to negotiate). Finally, we use the stock's *closing price* for S0.

## Bull spread using call options [TBU]
A bull call spread as a trading strategy consists in a long call with a low strike and a short call with a high strike, both with the same expiration date. The main goal is to generate a profit derived from sum of the payoff between the long and short calls. Three different positions of the stock price can be located and visualize if it generates a profit or a loss, if the stock price is below the low strike there is a profit of 0, when the stock price (ST) is above the high strike it generates a profit, therefore, if the stock price is between the low and high strike, it depends on the position of the stock and the cost of strategy value, in an mathematical form the value is the difference between the stock price and the low price plus the cost of strategy and it can generates a loss or a profit, which the loss is the cost of strategy.

First, we attempt to create a bull spread using call options for Tesla stock. 

### Strike and option prices for Tesla calls [TBU]
The stock price (*S0*) used is the same as in section 1, and the strike (*K1_cl, K2_cs*) and option (*P1_cl, P2_cs*) prices are the following:
(Reference current stock price shown in section 1 and add footnote with source for the images shown, and also include it in the bibliography)

```{r, echo = FALSE, message=FALSE, warning=FALSE}
Strike <- c("$710.00", "$725.00")
bid <- c("18.80", "11.25")
ask <- c("19.00", "11.55")
name <- c("Strike", "Bid", "Ask")

precios <-data_frame(Strike, bid, ask)
colnames(precios) <- name
kable(precios, caption = "Low and high strike values")
```

The variables that are going to be compute this scenario are:

- ST.seq: Range of the stock values at maturity date located in the x axis.

- $K_1$_cl: Strike price of the long put

- $K_2$_cs: Strike price of the short put

- $P_1$_cl: Price at the low strike

- $P_2$_cs: Price at the high strike

```{r}
# Set the range for the x axis (stock price at maturity) and define strike (K) and option (P) price for each call position used
ST.seq_bc <- seq(from = 650, to = 755, length.out = 105)

K1_cl <- 710.00
K2_cs <- 725.00
P1_cl <- 19.00
P2_cs <- 11.25

# Calculate the profit for each position
cl <- pmax(ST.seq_bc - K1_cl - P1_cl, -P1_cl) # long call profit function
cs <- pmin(K2_cs - ST.seq_bc +P2_cs, +P2_cs) # short call profit function
```

As a first and fundamental part of the Methodology, we start by setting the range of the sequence that will later be the x axis. As seen above the sequence will start from -PONER EL NUEMERO- to -PONER EL NUEMERO-, since the selected options have a strike price within the given range, the length out sets the length of the given sequence. It was decided to set that argument under the name "ST.seq", as ST stands for the final price of a given underlying asset, so that the expected profit will lay somewhere in between a point of a sequence. 

The next following part is to name the each of the basic arguments to use in the following steps, where:

-K1_cl: Strike price of the Long Call option
-K2_cs: Strike price of the Short Call option
-P1_cl: Price of the Long Call option
-P2_cl: Price of the Short Call option

On the very last part of this first chunk of the methodology to build a bull spread using call options, we have now, the most important part in order to create it, calculating the profit fr each of the positions. 

The first one, cl, which stands for Long Call, is set by the function pmax, the pmax function returns the parallel maxima of the input values. In this case we have two input vectors. The first one being ST.seq - K1_cl - P1_cl, in this part we are subtracting the strike price of the Long Call option and the price of the option form the sequence ST.seq, and the second argument being -P1_cl. The reason for the first vector, is to set the values of the of the sequence, while the second one, is for the Long Call option to start on the price of the option, and since it is a Long Call, the price is negative, it implicates a cash out in order to buy the option.

The second one, cs, which stands for Short Call, is set by the function pmin, the pmin function returns the parallel minima of the input values. In this case we have two input vectors. The first one being K2_cs - ST.seq + P2_cs, in this part we are setting the sequence ST.seq as a negative and adding the strike price of the Short Call option and the price of the option, and the second argument being +P2_cs. The reason for the first vector, is to set the values of the of the sequence, and since it is a Short Call it's supposed to be negative, that's the reason for ST.seq being set as a negative for the slope to go downwards, while the second one, is for the Short Call option to start on the price of the option, and since it is a Short Call, the price is positive, it implicates a cash in in order to buy the option.

As seen above, for the call option is used the function pmax, and on the short call option is used the pmin option, the reason for this, is for the slope to be in increasing or decreasing in the correct way,for the Long Call option, the slope is positive and increasing, and for the Short Call option the slope is negative and decreasing.

```{r}
# Graph the profit curve for each position
par(mfrow = c(1, 1), mai = c(0.7, 0.4, 0.4, 0.4))
par(pty = "s")
plot(ST.seq_bc, cl, type = "l", xlim = c(670, 730), ylim = c(-20, 3), lwd = 2, lty = 2, col = "orange",
     ylab = "Profit",
     xlab = "Stock price maturity")
abline(h = 0)
abline(v = K1_cl, lty = 2, col = "grey")
legend("topleft", legend = c("Long call"), col = "orange",
       lwd = 2, lty = 2, bg = "white", bty = "n")


par(pty = "s")
plot(ST.seq_bc, cs, type = "l",  xlim = c(690, 750), ylim = c(-20, 13), lwd = 2, lty = 2, col = "green",
     ylab = "Profit",
     xlab = "Stock price at maturity")
abline(h = 0)
abline(v = K2_cs, lty = 2, col = "grey")
legend("bottomleft", legend = c("Short call"), col = "green",
       lwd = 2, lty = 2, bg = "white", bty = "n")

```

The second chunk of the methodology is merely to visualize via a graph the movement of both options, the Long Call and the Short Call.

We started by calling the function par, which includes two different parameters, the first one, the parameter mfrow, is used to divide the graphic window into a matrix shape in order to store different graphics in each cells; the first value in parameter mfrow, which in this case is 2 is for the number of rows, and the second value, also 2, is for the number of columns. 
The second parameter of the function par is mai, this parameter is the one that is used to set the margins of the graphs measured in inches, in this case both top and side margins are set in 0.4 inches, while the inferior margin is set for 0.7 inches.

The next line of the code's chunk is again the function par, this time with the parameter pty, this parameter is used to specify the type of plot region that will be used, in this case the argument "s" generates a square plotting region.

The next and third line of the code's chunk is where the plotting begins, the function plot is a generic function used to graphic several items in R, in this case, the function will be used to plot profit curve of each of the call positions individually. 

The first plotted graphic is the graphic that shows the curve for the Long Call option, the first value of the option is the value of the x axis, in this case is ST.seq, which is the sequence set on the first chunk of the code, going from -PONER EL NUEMERO- to -PONER EL NUEMERO-, next, as the y axis, we set cl, which is the of the option that was also previously described in the first chunk of the methodology, the next value on the function is type, which in this time is defined as "l", this value refers to the type of plot to graph, and "l" stands for "lines". The nex value is ylim, this value helps set the range of the ylim to graph, in this case the ylim was set to go from -100 to 100, in order to visualize better the graph. The following part value of the fuction plot is lwd, which is the line width, for this case it's 2. The next two parts, ylab and xlab, are for labeling both the y and x axis in that order, for this plot our y axis is labeled as profit and the x axis as the stock price maturity. And Finally the last part for the graph of the Long Call curve is te legend, which, as seen in the graph in this case is goes for "Long Call".

For the following graph, that shows the Short Call, is basically built the same way as described above, we start again with the function part with the parameter pty setting the plot region again as "s".

Then we have the plot function for the graph, where in this case we set again  the x axis, as ST.seq, but for the y axis we set cs, which is the Short Call option describe in the chunk above.The next values of the function plot go the same as in the graph of the Long Call, the type of the plot was set as "l", we set the y axis on a range form -100 to 100 again, leave the line width as two, and put again the same labels for the y and x axis, lastly we put the legend on the graph as "Short Call" on the bottom left of the graph. 

The graphs help us visualize on a much graphic way the curve of the profit of both of the options, as seen and as it was briefly described above, the Long call profit curve starts at a negative, and once it hits the strike price, it starts increasing in values. For the Short Call options, it goes the other way, in this case the graph starts with a positive value and starts decreasing as the strike proce hits.

```{r}
# The bull spread is computed by adding the profit from each individual position 
profit.seq_bc <- ((cl + cs)) 

# Now we graph the bull spread profit created using call options
par(pty = "s")
plot(ST.seq_bc, profit.seq_bc, type ="l", xlim = c(690, 745), ylim = c(-20, 22), lwd = 2 , col = "red",
     main = "Figure 12.1a", xlab = "Stock price at maturity",
     ylab = "Profit")
lines(ST.seq_bc, cl, lty = 2, lwd = 2, col = "orange")
lines(ST.seq_bc, cs, lty = 2, lwd = 2, col = "green")
abline(v = 0)
abline(h = 0)
abline(v = K1_cl, lty = 2, col = "grey")
abline(v = K2_cs, lty = 2, col = "grey")
legend("topleft", cex = 0.8, legend = c("Long call", "Short call", "Bull call spread"),
col = c("orange", "green", "red"), lty = c(2, 2, 1),
lwd = 2, bg = "white")
```

On the next part we calculate the bull spread, naming it "profit.seq_bc", the way of doing it was by adding the profits of each of the call options, the reason for this has an ----geometric---- reason as well as a mathematical reason, we intend to add both of the sequences in order to leave the sequence of the profit that both profit would have as a one.

The last part of the Chunk is the actual plot of the whole bull spread profit, as in the same methodology of the last two plots, we start by setting the pty parameter as "s", and then proceed to plot the bull spread profit by setting the x axis as ST.seq and the y axis as profit.seq_bc. After having the plot for the profit.seq_bc ready, we proceed to add the lines of both of the call options, setting the Long Call option as an orange color and the Short Call option as the blue line.

As it can be seen on the graph above the profit range that our bull spread could have goes from -7.75, since that's the price of buying both of our call options, to a maximum profit of 7.25.

The following table display the profit or losses values according to the position of the stock price:

```{r, echo = FALSE}
code_bc <- c("-7.75", "-7.74 to 7.24", "7.25")
pricer_bc <- c("ST <= K_1", "K_1 < ST < K_2", "ST >= K_2")
name_bc <- c("Stock Price", "Profit or Loss")
range_bc <- data_frame(pricer_bc, code_bc)
colnames(range_bc) <- name_bc
kable(range_bc, caption = "Bull call spread profit range")
```

Commentary about kable... [TBU]

## Bull spread using put options [TBU]
Another trading strategy with bull spread is using put options, this strategy operates two positions, buying a put option with a low strike price and writing a put option with a high strike price, both having the same maturity date. Bull put spread consists in expecting that the strike price increases and consequently generates a gain to the investor.

Generating a profit is the objective and according to the bull put spread conditions there are the positions where can create a profit or loss depending on the location of the stock price also known as ST at the maturity date, when the ST is above the high strike generates a profit to the investor, otherwise if the ST is lower than the low strike it will cause a loss, and the third case is when the st is inside the spread (between the strikes) the profit or loss will depend on the result of the low strike plus the cost of strategy and can cause either a profit or a loss depending the case.

This section continues using Tesla stock as a form of example but applying the theory of the bull spread strategy using put options and computing the results for a better visualization and comprehension. First of all, it is important to determine the values that are going to be computing and subsequently the variables. 

yahoo finance

The image )2.3) represents the value of stock prices during a period of time and it’s bid and ask price 

The following table shows the selected values with their corresponding bod and ask price. In this case, the selected values have a difference of 50 dollars between them, which is going to be the selected spread, according to the theory when you are buying an option (low strike) the value that you are paying is going to be the ask price and when you are selling an option (high strike) the price at you are going to sell is the bidding price.

### Strike and option prices for Tesla puts [TBU]
Reference current stock price shown in section 1
The strike (*K1_pl, K2_ps*) and option (*P1_pl, P2_ps*) prices for the puts used are the following:
(Reference current stock price shown in section 1 and add footnote with source for the images shown, and also include it in the bibliography)

```{r, echo = FALSE, message=FALSE, warning=FALSE}
Strike <- c("$705.00", "$720.00")
bid <- c("14.80", "21.50")
ask <- c("15.20", "22.05")
name <- c("Strike", "Bid", "Ask")

precios <-data_frame(Strike, bid, ask)
colnames(precios) <- name
kable(precios, caption = "Low and high strike values")
``` 

The variables that are going to be compute this scenario are:

- ST.seq: Range of the stock values at maturity date located in the x axis.

- $K_1$_pl: Strike price of the long put

- $K_2$_pc: Strike price of the short put

- $P_1$_pl: Price at the low strike

- $P_2$_pc: Price at the high strike

```{r}
# Set the range for the x axis (stock price at maturity) and define strike (K) and option (P) price for each put position used
ST.seq_bp <- seq(from = 650, to = 755, length.out = 105) 

K1_pl <- 705
K2_ps <- 720
P1_pl <- 15.20
P2_ps <- 21.50

# Calculate the profit for each position
pl <- pmax(K1_pl - P1_pl - ST.seq_bp, -P1_pl) # long put profit function
ps <- pmin(ST.seq_bp - K2_ps +P2_ps, +P2_ps) # short put profit function
```

Having the procedure of long and short put allows to perform a plot to have a better visualization of each option in an individual form.

```{r}
# Graph the profit curve for each position
par(mfrow = c(1, 1), mai = c(0.7, 0.4, 0.4, 0.4))
par(pty = "s")
plot(ST.seq_bp, pl, type = "l", xlim = c(680, 740), ylim = c(-16, 10), lwd = 2, lty =2 , col = "orange",
     ylab = "Profit",
     xlab = "Stock price at maturity")
abline(h = 0)
abline(v = K1_pl, lty = 2, col = "grey")
legend("topright", legend = c("Long put"), col = "orange",
       lwd = 2, lty =2, bg = "white", bty = "n")

par(pty = "s")
plot(ST.seq_bp, ps, type = "l", xlim = c(690, 740), ylim = c(-10, 23), lwd = 2, lty =2 , col = "green",
     ylab = "Profit",
     xlab = "Stock price at maturity")
abline(h = 0)
abline(v = K2_ps, lty = 2, col = "grey")
legend("bottomright", legend = c("Short put"),  col = "green",
       lwd = 2, lty =2, bg = "white", bty = "n")
```

The long put plot is located on the left side and shows its behavior were we can visualize how if the stock price is lower the payoff is higher and its going downwards until reaching the low strike, this means that there is a profit for the investor while the stock price is lower than the strike and if it is above the strike price there will be a loss due to the stock price is higher than the option price that its establish. On the other hand, the short put in on the right side of the image, the behavior of the slope demonstrate that while the stock price is higher there is going to be a profit until reaching the strike will continue increasing.

After analyzing the behavior of each option, we can start to compute the profit deviated from the bull put spread strategy option, the next code shows the mentioned calculation, where *S0* means the current stock price. Furthermore, there is a new data assign called *profit.seq.bp* which is the bull spread computed by the addition of the long put and the short put previously calculated.

```{r}
# The bull spread is computed by adding the profit from each individual position
profit.seq_bp <- ((pl + ps)) 

# Now we graph the bull spread profit created using put options
par(pty = "s")
plot(ST.seq_bp, profit.seq_bp, type ="l", xlim = c(690, 735), ylim = c(-15, 25), lwd = 2, col = "blue",
     main = "Figure 12.1a", xlab = "Stock price at maturity",
     ylab = "Profit")
lines(ST.seq_bp, pl, lty = 2, lwd = 2, col = "orange")
lines(ST.seq_bp, ps, lty = 2, lwd = 2, col = "green")
abline(v = 0)
abline(h = 0)
abline(v = K1_pl, lty = 2, col = "grey")
abline(v = K2_ps, lty = 2, col = "grey")
legend("topleft", cex = 0.8, legend = c("Long call", "Short call", "Bull put spread"),
col = c("orange", "green", "blue"), lty = c(2, 2, 1),
lwd = 2, bg = "white")
```

The put spread plot specify the profit of the investor depending on the position of the stock price in the x axis, in this case the range of the stock price is from 600 to 800 where we can visualize the behavior of the bull spread profit, the vertical gray dashed lines specify the lower and the high strikes, the color dashed lines represents the put options and the red line is the profit at certain stock price. As we mentioned before in the theory, with this graph we can analyze and visualize if the stock price is higher than the low strike it will generate a profit and is continuously increasing until it reaches the high strike, moreover if the stock price is under the low strike, there is a loss.

The following table display the profit or losses values according to the position of the stock price:

```{r, echo = FALSE}
code_bp <- c("-2.20", "-2.19, to 47.79", "47.80")
pricer_bp <- c("ST <= K_1", "K_1 < ST < K_2", "ST >= K_2")
name_bp <- c("Stock Price", "Profit or Loss")
range_bp <- data_frame(pricer_bp, code_bp)
colnames(range_bp) <- name_bp
kable(range_bp, caption = "Bull put spread profit range")
```

Commentary about kable... [TBU]

## Bull spread comparison [TBU]
The curves shown above demonstrate it is possible to design a profitable trading strategy by building a bull spread using both call and put options. Now, we show both spread curves in one same plot in order to easily visualize how similar they are. In order to do this, we use the same *par*, *plot* and *lines* functions used on the other curves shown along the paper. Note we use the *ST.seq* variables for the x axis and a *vector* for the y range (-100, 150). Finally, we make the call spread curve red and the put spread curve blue. See below:

```{r}
# Now we graph the bull spread profit created using put options
par(pty = "s")
plot(ST.seq_bc, profit.seq_bc, type ="l", xlim = c(690, 735), ylim = c(-10, 9), lwd = 2, col = "red",
     main = "Bull spread comparison", xlab = "Stock price at maturity",
     ylab = "Profit")
lines(ST.seq_bp, profit.seq_bp, type ="l", xlim = c(690, 735), ylim = c(-10, 10), lwd = 2, col = "blue")
legend("topleft", cex = 0.8, legend = c("Bull call spread", "Bull put spread"),
col = c("red", "blue"), lty = c(1,1),
lwd = 2, bg = "white")
```

As seen on the plot above, both spread-trading strategies have a very similar level of profitability, as their curves intersect on some points...[TBU]

# Conclusion [TBU]

# Honor pledge
We, Rodrigo, Valerie, Marcela, María Fernanda, Elsy & Rafael, hereby declare we have worked on this assignment while maintaining academic integrity.

# Bibliography [TBU]
Hull
Lozano, M.
Yahoo Finance
